A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-01-15 paper/draft.tex
expand context and data

This paper contributes to literature on clientelism that outlines the political logic of patronage allocations. The contribution is twofold: 1) first, I find that patronage is an elite-elite game, generating a form of patronage that is distinct from the politician-voter nexus that has been traditionally the focus of extant literature on clientelism \citet{diaz2016political} and 2) I find that the benefits of patronage go beyond simply an electoral pay-off. Instead, what patronage accomplishes is securing access to economic resources that can be tapped into once employment is offered to wealthy patrons, similar in vein to the theoretical findings by \citet{robinson2013political}. Patronage ties the hands of parties and patrons, in effect capturing the benefits of public sector employment to finance efforts towards party-building.

@@

>>>>

This paper contributes to literature on clientelism that outlines the political logic of patronage allocations. The contribution is twofold: 1) first, I find that patronage is a patron-elite game, generating a form of patronage that is distinct from the politician-voter nexus that has been traditionally the focus of extant literature on clientelism \citet{stokes2013brokers, diaz2016political} and 2) I find that the benefits of patronage go beyond simply an electoral pay-off. Instead, what patronage accomplishes is securing access to economic resources that can be tapped into once employment is offered to wealthy patrons, similar in spirit to the theoretical findings by \citet{robinson2013political}. Patronage binds parties and patrons, in effect capturing the benefits of public sector employment to finance efforts towards party-building.

The paper is structured as follows. Section \ref{sec:context} provides context for the data and the hiring process for local bureaucracies, while section \ref{sec:descriptive} provides some descirptive statistics outlining the differences between partisan and non-partisan members. Section \ref{sec:empirical} outlines the empirical strategy. Section \ref{sec:conclusion} concludes.

@@

\subsection{Partisan affiliation in Brazil}

Partisanship in Brazil is voluntary, with over 11 percent of registered voters affiliated to a party \citep{}

There is an ongoing debate on the strength of partisan ties in Brazil. On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citet{desposato2006parties,ames2002deadlock}. On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}. Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- which disproportionately

\section{Descriptive Statistics}
\label{sec:descriptive}
